### 2025-07-01 paper.tex
Write more

\documentclass[acmsmall, screen, nonacm]{acmart}

\usepackage{listings}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{fontawesome5}
\usepackage{mathpartir}
\usepackage{sourcecodepro}
\usepackage{amsthm}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\lstset{
  basicstyle=\ttfamily\small,
  columns=fullflexible,
  keepspaces=true
}

\lstdefinestyle{ocaml}{
  keywordstyle=\color{blue},
  emphstyle=\color{violet},
  language=[Objective]Caml,
  morekeywords={effect, exclave},
  escapeinside={(*@}{@*)},
  literate=
    {[]}{$\glob$}{1}
}

\lstdefinestyle{haskell}{
  keywordstyle=\color{blue},
  emphstyle=\color{violet},
  language=Haskell
}

\newcommand{\nat}{\mathbb{N}}
\newcommand{\glob}{\mathop{\Box}}
\newcommand{\cat}{\mathbb{C}}
\newcommand{\op}{\mathtt{op}}
\newcommand{\catop}{\cat^{\op}}
\newcommand{\id}[1]{\mathrm{id}_{#1}}
\newcommand{\setc}{\mathrm{Set}}
\newcommand{\psh}{\widehat{\cat}}
\newcommand{\terminal}{\top}
\newcommand{\initial}{\bot}
\newcommand{\lawc}{\mathrm{Law}}
\newcommand{\lawcop}{\lawc^{\mathtt{op}}}
\newcommand{\pshlawcop}{\widehat{\lawcop}}
\newcommand{\freef}{Free}
\newcommand{\constfname}{\Omega}
\newcommand{\constf}[1]{\constfname #1}
\newcommand{\yonedaname}{y}
\newcommand{\yoneda}[1]{y(#1)}
\newcommand{\homset}[3]{\mathrm{Hom}_{#1}(#2, #3)}
\newcommand{\arrow}[2]{#1 \Rightarrow #2}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\reflectname}{\Uparrow}
\newcommand{\reflectraw}[1]{\mathop{\reflectname_{#1}}}
\newcommand{\reflect}[3]{\reflectraw{#1}(#2, #3)}
\newcommand{\reflectc}[2]{\mathop{\reflectname_{#1 #2}}}
\newcommand{\performraw}[2]{\mathop{\reflectname(#1(#2))}}
\newcommand{\perform}[5]{\performraw{#1}{#2}(#3, #4. #5)}
\newcommand{\reifyname}{\Downarrow}
\newcommand{\reifyraw}[1]{\mathop{\reifyname_{#1}}}
\newcommand{\reify}[3]{\reifyraw{#1}(#2.\,#3)}
\newcommand{\reifyc}[2]{\mathop{\reifyname_{#1 #2}}}
\newcommand{\outl}[1]{\mathop{\mathrm{out}_L} #1}
\newcommand{\outr}[1]{\mathop{\mathrm{out}_R} #1}
\newcommand{\mon}{\mathcal{T}}
\newcommand{\ind}[1]{W_{#1}}
\newcommand{\indintro}[4]{#2_{#1}(#3, #4)}
\newcommand{\retname}{\mathrm{ret}}
\newcommand{\ret}[2]{\mathop{\retname_{#1}}(#2)}
\newcommand{\indelim}[3]{\mathop{\mathrm{rec}_{#1}} #2 \{#3\}}
\newcommand{\indretcase}[2]{\retname(#1) \Rightarrow #2}
\newcommand{\indcase}[4]{#1(#2, #3) \Rightarrow #4}
\newcommand{\seq}{\,;\,}
\newcommand{\fst}{\mathop{\pi_1}}
\newcommand{\snd}{\mathop{\pi_2}}
\newcommand{\inl}[1]{\mathop{\mathrm{in}_L} #1}
\newcommand{\inr}[1]{\mathop{\mathrm{in}_R} #1}
\newcommand{\case}[5]{\mathop{\mathrm{case}} #1 \,\{\,\inl{#2} \Rightarrow #3 \seq \inr{#4} \Rightarrow #5 \,\}}
\newcommand{\abs}[3]{\mathop{\lambda}(#1 \types #2).\,#3}
\newcommand{\app}[2]{#1\,#2}
\newcommand{\types}{\mathrel{:}}
\newcommand{\gtypes}{\mathrel{:_{\Box}}}
\newcommand{\cempty}{\varepsilon}
\newcommand{\ccons}[2]{#1;\,#2}
\newcommand{\lbind}[3]{\ccons{#1}{#2\types#3}}
\newcommand{\gbind}[3]{\ccons{#1}{#2\gtypes#3}}
\newcommand{\cdiv}[1]{#1/\Box}
\newcommand{\return}[1]{\mathop{\mathrm{return}} #1}
\newcommand{\letv}[3]{\mathop{\mathrm{let}} #1 = #2 \mathop{\mathrm{in}} #3}
\newcommand{\gbox}[1]{\mathop{\mathrm{box}} #1}
\newcommand{\gunbox}[4]{\mathop{\mathrm{let}} (\gbox #1) = #2 \mathop{\mathrm{in}} #3}
\newcommand{\grun}[1]{\mathop{\mathrm{run}}(#1)}
\newcommand{\sbst}[3]{#1[#2\backslash#3]}
\newcommand{\turnv}{\mathrel{\vdash_V}}
\newcommand{\turnc}{\mathrel{\vdash_C}}
\newcommand{\evmap}{\mathop{ev}}
\newcommand{\currymap}{\mathop{\lambda}}
\newcommand{\reducesto}{\rightsquigarrow}

\newcommand{\todo}[1]{{\color{red}TODO: #1}}

\title{Locality and Effect Reflection}
\author{Leo White}
\affiliation{\institution{Jane Street}\country{UK}}
\date{}
\setcopyright{none}

\begin{document}

\begin{abstract}
  Algebraic effects and handlers allow the programmer to extend their
  language's \emph{ambient monad} \textbf{--} the monad describing which
  effects ordinary functions can perform \textbf{--} with additional
  effectful operations.

  Locality captures the general idea of a value depending on some notion
  of ``location''. We can represent it in a type system using a
  comonadic \emph{global modality} $\glob$ that tracks which values do
  not depend on their location. A particularly useful notion of
  ``location'' is the ambient monad.

  We introduce an alternative presentation of algebraic effects and
  handlers, \emph{effect reflection}, that provides, for any algebraic
  signature $\Sigma$, an isomorphism:
  \begin{equation*}
    \ind{\Sigma}(\glob \tau) \qquad \cong \qquad \yoneda{\Sigma} \rightarrow \glob \tau
  \end{equation*}
  where $\ind{\Sigma}$ is an inductive type representing the free
  algebra of $\Sigma$, and $\yoneda{\Sigma}$ is a type representing how
  to inject terms of $\Sigma$ into the current ambient
  monad. $\ind{\Sigma}$ is sometimes called the \emph{freer monad} of
  $\Sigma$. This presentation falls out naturally from a simple
  denotational semantics for algebraic effects in terms of presheaves
  over bounded Lawvere theories.

  Effect reflection lets you work with any algebraic monad using an
  ordinary function type, allowing effectful code to be written in
  direct style. This provides all the benefits of typed algebraic
  effects and handlers without requiring an effect system. We
  demonstrate effect reflection by implementing it as a library in
  OxCaml\cite{lorenzen2024oxidizing,} and working through some classic
  examples of programming with effects.
\end{abstract}

\maketitle

\section{Reflection for algebraic effects}

\paragraph{Algebraic effects} Algebraic effects and handlers are an
increasingly popular approach to programming with effects. Handlers
allow the programmer to add additional effects to the \emph{ambient
  monad} of their language -- the monad describing which effects an
ordinary function can perform -- for some region of their program.

For example, in OCaml, one can write:
\begin{lstlisting}[style=ocaml]
type 'a Effect.t +=
  | Get : int Effect.t
  | Set : int -> unit Effect.t

let handle_state (f : unit -> 'a) : 'a =
  match f () with
  | x -> fun _ -> x
  | effect Get, k -> fun s -> continue k s s
  | effect Set s', k -> fun _ -> continue k () s'
\end{lstlisting}
and for the duration of the \lstinline[style=ocaml]{f ()} function call
the ambient monad is extended with two additional effectful operations:
\begin{lstlisting}[style=ocaml]
let get () : int = perform Get
let set (i : int) : unit = perform (Set i)
\end{lstlisting}

Note that, in OCaml, nothing prevents attempting to perform an algebraic
effect operation that is not supported by the current ambient monad,
resulting in an \lstinline[style=ocaml]{Unhandled} exception being
raised. Other languages, e.g. Koka\cite{??}, use an effect system to
statically ensure operations are only performed when they are available.

\paragraph{Locality and effect reflection}
Locality captures the general idea of a value depending on some notion
of ``location''. We can represent it in a type system using a comonadic
\emph{global modality} $\glob$ that tracks which values do not depend on
their location. A particularly useful notion of ``location'' is the
ambient monad.

Treating the ambient monad as a notion of location allows a language to
provide an isomorphism:
\begin{equation*}
  \mathop{M}(\glob X) \qquad \cong \qquad Y \mathbin{\rightarrow} \glob X
\end{equation*}
where $M$ is any algebraic monad, $Y$ is a specific type related to that
monad, and $\glob$ is a comonadic global modality. We call this
isomorphism \emph{effect reflection}.

\todo{Restrict the above claim to signatures}

The ability to use the ordinary arrow type
$Y \mathbin{\rightarrow} \glob X$ in place of the monad
$\mathop{M}(\glob X)$ allows us to conveniently write effectful
code in direct style, just as algebraic effects do.

\paragraph{Effect reflection in OxCaml}
OxCaml\cite{lorenzen2024oxidizing} provides type system support for
tracking locality and runtime support for algebraic effects. For
backwards compatibility with OCaml, an unadorned arrow type
\lstinline[style=ocaml]{t -> s} is treated as if the parameter and
return types were actually \lstinline[style=ocaml]{[]t} and
\lstinline[style=ocaml]{[]s}. To get parameters or results that are not
implicitly global, we write
e.g. \lstinline[style=ocaml]{t @ local -> s @ local}.

Using this support, with can provide effect reflection as a simple API
that we can use for programming with effects.  For example, we can use
it to write an effect handler equivalent to the OCaml one above:
\begin{lstlisting}[style=ocaml]
module State = struct
  type 'a t =
    | Get : int t
    | Set : int -> unit t
end

module State_eff = Reflection(State)

let handle_state (f : Handler(State).t @ local -> 'a) (i : int) : 'a =
  let rec loop s : 'a Term(State).t @ local -> 'a  = function
    | Return x -> x
    | Op(Get, k) -> loop s (k s)
    | Op(Set s', k) -> loop s' (k ())
  in
  loop i (State_eff.reify_local f)
\end{lstlisting}
which calls \lstinline[style=ocaml]{f} and passes it a local value of
type \lstinline[style=ocaml]{Handler(State).t}. For the duration of the
call to \lstinline[style=ocaml]{f} the ambient monad is extended with
two additional effectful operations, which can be used via reflection:
\begin{lstlisting}[style=ocaml]
let get (h : Handler(State).t @ local) : int =
  State_eff.perform Get h

let set (h : Handler(State).t @ local) (i : int) : unit =
  State_eff.perform (Set i) h
\end{lstlisting}

Note that the locality of the handler value ensures that these
operations can only be performed when they are available in the ambient
monad. This provides the same guarantees as using an effect system.

Effect reflection is closely related to \emph{lexical effect handlers} in
systems such as Effeckt\cite{??}. Lexical effect handlers have a value that
represents the effect handler, but these values are second class and
require their own special typing rules.

Effect reflection is also closely related to \emph{named effect
  handlers}\cite{??}. Named effect handlers use higher-rank polymorphism
to emulate locality tracking in much the same way as the
\lstinline[style=haskell]{St} monad in Haskell. Using higher-rank
polymorphism prevents global type inference, requiring complex inference
that extends beyond classic Hindly-Milner, and demanding a higher
annotation burden on the programmer. \todo{Mention the explicit monad
that named effect handlers presumably need.}

Effect reflection is a variation on monadic reflection\cite{??}.
\todo{Make this relationship precise.}

\paragraph{Denotational semantics} The design of effect reflection falls
out naturally from a simple denotational semantics of algebraic
effects. The underlying idea is to index the semantics by the ambient
monad. Rather than the semantics of a type being a set it is instead a
function from ambient monads to set: if you tell it what the ambient
monad is it can tell you what its set of values is.

The semantics is defined in terms of presheaves of the opposite of the
category of \emph{bounded Lawvere theories}. Bounded Lawvere theories
are a categorical representation of algebraic theories: we represent the
ambient algebraic monad by the corresponding algebraic theory. The key
constructs of effect reflection are then defined as follows:

\begin{itemize}
\item The ambient monad can be defined as a monad $\mon$ on such
  presheaves, which maps each algebraic theory to its free algebra. We use
  $\mon$ to define the semantics of computations.

\item The global modality $\glob$ on such presheaves is defined such
  that $\glob(P)(L) = P(\initial)$. In other words, for a given presheaf
  $P$ it gives the constant presheaf that always ignores the ambient
  monad and instead returns $P$ at $\initial$, where $\initial$ is the
  algebraic theory with no operations or equations. This allows us to
  restrict a type to its terms that do not depend on the ambient monad.

\item The semantics of our handler types $Y$ are given by the Yoneda
  embedding, and our effect reflection isomorphism becomes a simple
  consequence of the Yoneda lemma.
\end{itemize}

These presheaf semantics are closely related to presheaf semantics for
higher-order abstract syntax\cite{??}.

\paragraph{Structure of the paper} We proceed as follows:
\begin{itemize}
\item In Section \ref{sec:pracice}, we give an overview of effect reflection in
  practice in OxCaml
\item In Section \ref{sec:semantics}, we describe the constructions for a
  novel denotational semantics of algebraic effects
\item In Section \ref{sec:language}, we define a simply typed calculus featuring
  effect reflection along with its denotational semantics
\item In Section \ref{sec:implementation}, we give an implementation of effect
  reflection in OxCaml in terms of the untracked algebraic effects it
  inherits from OCaml.
\item In Section \ref{sec:related-work}, we discuss related work, and in
  Section \ref{sec:future-work}, we discuss future work.
\end{itemize}

\section{Effect reflection in practice}
\label{sec:pracice}

We can provide effect reflection as a simple library API in OxCaml. In
this section we describe this API and show how to use it to write a
number of typical examples of algebraic effects.

\subsection{Locality in OxCaml}

The effect reflection API relies on OxCaml's support for tracking
locality in types, so let's start with an overview of that support.

Values in OxCaml are associated with both a type -- which describe how
values can be introduced and eliminated -- and a \emph{mode} -- which
track properties related to operations other than introduction and
elimination. Modes are a product of a number of different axes, tracking
a variety of different properties, but here we are only interested in
one of them: \emph{locality}. Ignoring the other axes, a mode is either
\lstinline[style=ocaml]{local} or \lstinline[style=ocaml]{global}.

Locality tracks the ability of a value to leave \emph{regions}.  Regions
describe some period of time during a program's execution. Every
function call defines a corresponding region that, by default, lasts for
the entire extent of the call. This means that, by default,
\lstinline[style=ocaml]{local} values cannot be returned from a
function.
\begin{lstlisting}[style=ocaml]
let leak_local x =
  let (o @ local) = Some x in
  o
(*@ \textcolor{red}{Error: o} @*)
       (*@ \textcolor{red}{\textasciicircum} @*)
(*@ \textcolor{red}{This value is local and would escape its region.} @*)
\end{lstlisting}

To support functions that return local values, the region of a function
can be ended early using the \lstinline[style=ocaml]{exclave} construct.
\begin{lstlisting}[style=ocaml]
let return_local x =
  exclave
    (let (o @ local) = Some x in
     o)
\end{lstlisting}
\lstinline[style=ocaml]{exclave e} can only appear in tail position
within a function.  It ends the function's region early and then
executes the expression
\lstinline[style=ocaml]{e}. \lstinline[style=ocaml]{e} is essentially
executed within the region of the caller of the function and any local
values created within it can be safely returned to that caller.

The place where modes appear within the type algebra is on arrow
types. A function being \lstinline[style=ocaml]{global} doesn't tell you
that the values it accepts or returns should be
\lstinline[style=ocaml]{global} too, so instead the arrow type is
labelled with two modes: one for the argument and one for the return
value.
\begin{lstlisting}[style=ocaml]
val return_local : 'a @ local -> 'a option @ local
\end{lstlisting}
For convenience, and backwards compatibility with OCaml, either mode can
be omitted, in which case it defaults to
\lstinline[style=ocaml]{global}.

By default, modes are deep: a local \lstinline[style=ocaml]{string list}
is a local list of local strings. However, the depth that a mode applies
can be controlled via \emph{modalities}. Record fields can be annotated
with \lstinline[style=ocaml]{global} modality to indicate that the
contents of the field are global even when the surrounding record is
local. For example:
\begin{lstlisting}[style=ocaml]
type 'a gbl = { g : 'a @@ global }
\end{lstlisting}
gives a record whose only field has the \lstinline[style=ocaml]{global}
modality. The contents of the \lstinline[style=ocaml]{g} field is always
global, for example:
\begin{lstlisting}[style=ocaml]
let project_global (r @ local) =
  r.g
\end{lstlisting}
gives a function of type:
\begin{lstlisting}[style=ocaml]
val project_global : 'a gbl @ local -> 'a @ global
\end{lstlisting}
Variant constructor arguments can also be annotated with modalities.

Note that, whilst there is a \lstinline[style=ocaml]{global} modality,
there is no \lstinline[style=ocaml]{local} modality. It does not make
sense for a value which has the ability to leave a region to contain a
value which doesn't, as when the first value leaves a region it will
implicitly cause the second one to leave that region as well. This means
the \lstinline[style=ocaml]{global} mode is always deep.

\subsection{Effect reflection interface}
\label{sec:interface}

The types and operations of our effect reflection API are parameterized
by type constructors representing the operations of an algebraic
effect. In OxCaml, such higher-kinded interfaces must be built using
functors, so we start with a simple module type
\lstinline[style=ocaml]{Op} to act as the parameter type for these
functors:
\begin{lstlisting}[style=ocaml]
module type Op = sig type 'a t end
\end{lstlisting}
The fundamental types of the API are the \lstinline[style=ocaml]{Term}
and \lstinline[style=ocaml]{Handler} types.
\begin{lstlisting}[style=ocaml]
module Term (O : Op) : sig
  type 'a t =
    | Return : 'a @@ global -> 'a t
    | Op : 'r O.t @@ global * ('r -> 'a t) -> 'a t
end

module Handler(O : Op) : sig
  type t
end
\end{lstlisting}
For an effect signature \lstinline[style=ocaml]{O},
\lstinline[style=ocaml]{'a Term(O).t} is the type of terms in the free
algebra of the corresponding effect -- otherwise known as the freer
monad\cite{??}. \lstinline[style=ocaml]{Handler(O).t} is the type of
mappings from the operations of \lstinline[style=ocaml]{O} into the
operations of the current ambient monad. Alternatively,
\lstinline[style=ocaml]{Handler(O).t} can be described as representing
algebraic effect handlers that are handling the current computation --
hence the name.

The main piece of the API is a family of isomorphisms:
\begin{equation*}
  \text{\lstinline[style=ocaml]{'a Term(O).t}} \qquad
  \cong \qquad \text{\lstinline[style=ocaml]{Handler(O).t @ local -> 'a}}
\end{equation*}
Due to the absence of mode polymorphism in OxCaml, we expose this
isomorphism as three functions: \lstinline[style=ocaml]{reify} and
\lstinline[style=ocaml]{reify_local} which implement the left-to-right
direction at the \lstinline[style=ocaml]{global} and
\lstinline[style=ocaml]{local} modes respectively; and
\lstinline[style=ocaml]{perform}, which is equivalent to the
right-to-left direction but makes for a more convenient primitive in
practice. \lstinline[style=ocaml]{perform} amounts to reflection of
individual operations as opposed to full terms.
\begin{lstlisting}[style=ocaml]
module Reflection (O : Op) : sig
  val reify : (Handler(O).t @ local -> 'a) -> 'a Term(O).t

  val reify_local :
    (Handler(O).t @ local -> 'a) @ local -> 'a Term(O).t @ local

  val perform : 'a O.t -> (Handler(O).t @ local -> 'a)
end
\end{lstlisting}
Using \lstinline[style=ocaml]{perform} we can implement reflection of
terms as:
\begin{lstlisting}[style=ocaml]
let rec reflect t h =
  match t with
  | Return x -> x
  | Op(op, k) -> reflect (k (perform op h)) h

let rec reflect_local (t @ local) h =
  match t with
  | Return x -> x
  | Op(op, k) -> reflect_local (k (perform op h)) h
\end{lstlisting}
with types:
\begin{lstlisting}[style=ocaml]
val reflect : 'a Term(O).t -> (Handler(O).t @ local -> 'a)

val reflect_local :
   'a Term(O).t @ local -> (Handler(O).t @ local -> 'a) @ local
\end{lstlisting}

\subsection{Example: State}

As shown in the introduction we can implement an effect handler for
state using this API. We start by defining the operations of the effect:
\begin{lstlisting}[style=ocaml]
module State = struct
  type 'a t =
    | Get : int t
    | Set : int -> unit t
end
\end{lstlisting}
Then we instantiate the reflection isomorphism for that effect and use
it to write a handler:
\begin{lstlisting}[style=ocaml]
module State_eff = Reflection(State)

let handle_state f i =
  let rec loop s = function
    | Term(State).Return x -> x
    | Term(State).Op(Get, k) -> loop s (k s)
    | Term(State).Op(Set s', k) -> loop s' (k ())
  in
  loop i (State_eff.reify_local f)
\end{lstlisting}
Finally we add an alias for the handler type and some convenient
wrappers for the generic effects:
\begin{lstlisting}[style=ocaml]
type handler = Handler(State).t

let get h =
  State_eff.perform Get h

let set h i =
  State_eff.perform (Set i) h
\end{lstlisting}
Giving us the following interface:
\begin{lstlisting}[style=ocaml]
type handler

val handle_state : (handler @ local -> 'a) @ local -> int -> 'a
val get : handler @ local -> int
val set : handler @ local -> int -> unit
\end{lstlisting}
Note that we are able to use \lstinline[style=ocaml]{reify_local} to allow
the function passed to \lstinline[style=ocaml]{handle_state} to be local.

This shows how to implement state using effect reflection, but note that
the same interface can also be implemented direcly using mutable state
and locality:
\begin{lstlisting}[style=ocaml]
type handler = int ref

let handle_state f i = f (ref i)
let get r = !r
let set r i = r := i
\end{lstlisting}
This implementation gives the same behaviour as the one built using
effect reflection. Crucially it is just as easy to reason about, which
in turn means it is just as easy to reason about as the state monad. In
some sense, the improved reasoning ability of the state monad over
ordinary mutable state is precisely captured by the notion of locality.

This transformation from effect relection to local mutable state works
for any \emph{tail-resumptive} effect handler.

\subsection{Lightweight effect polymorphism}

The following function uses state to calculate the total of a list of
integers:
\begin{lstlisting}[style=ocaml]
let total l =
  handle_state (fun h ->
    List.iter (fun x -> set h (get h + x)) l;
    get h)
\end{lstlisting}
This relies on the locality of the function parameter of
\lstinline[style=ocaml]{List.iter}:
\begin{lstlisting}[style=ocaml]
val iter : ('a -> unit) @ local -> 'a list -> unit
\end{lstlisting}
The \lstinline[style=ocaml]{local} annotation on the parameter ensures
that any function passed will not escape the current region. This allows
us to pass in a closure that closes over the
\lstinline[style=ocaml]{local} handler \lstinline[style=ocaml]{h} and
use it to perform some effects.

This ability for higher-order functions like
\lstinline[style=ocaml]{List.iter} to be reused with different ambient
monads achieves the ``lightweight effect polymorphism'' described by ??
et al.\cite{??} without requiring a separate notion of second-class
function.

\subsection{Example: Generators}

A more realistic use of effect reflection is a \emph{generator}
effect. This effect has a single operation:
\begin{lstlisting}[style=ocaml]
module Gen = struct
  type 'a t =
    | Gen : int -> unit t
end
\end{lstlisting}
Along with a handler that turns computations using
\lstinline[style=ocaml]{Gen} into streams of integers.
\begin{lstlisting}[style=ocaml]
module Gen_eff = Reflection(Gen)

type ints =
  | Finished
  | More of int * (unit -> ints)

let handle_gen f =
  let rec loop = function
    | Term(Gen).Return () -> Finished
    | Term(Gen).Op(Gen i, k) -> More(i, k)
  in
  loop (Gen_eff.reify f)
\end{lstlisting}

Unlike the state handler, here we must use
\lstinline[style=ocaml]{reify} rather than
\lstinline[style=ocaml]{reify_local} because our intention is that the
resulting stream of integers be a global value that can escape the
current region. Since it closes over the continuation
\lstinline[style=ocaml]{k}, we must use \lstinline[style=ocaml]{reify}
and require that \lstinline[style=ocaml]{f} be global:
\begin{lstlisting}[style=ocaml]
val handle_gen : (Handler(Gen).t @ local -> unit) -> ints
\end{lstlisting}

\subsection{Example: Asynchronous I/O}

Perhaps the most common use case for algebraic effects is for
concurrency and asynchronous I/O. Implementing asynchronous I/O from
scratch is beyond the scope of this paper, but we can build a simple
version on top of an existing monadic concurrency library. Assuming we
have a monadic future type \lstinline[style=ocaml]{'a Deferred.t} we can
write concurrent code in direct style by using an effect with a single
\lstinline[style=ocaml]{Await} operation that waits for a future to be
completed:
\begin{lstlisting}[style=ocaml]
module Await : sig
  type 'a t =
    | Await : 'a Deferred.t -> 'a t
end
\end{lstlisting}

\begin{lstlisting}[style=ocaml]
module Await_eff = Reflection(Await)

let handle_await f =
  let rec loop = function
    | Term(Await).Return x -> Deferred.return x
    | Term(Await).Op(Await d, k) ->
        Deferred.bind d (fun x -> loop (k x))
  in
  loop (Await_eff.reify f)

let await h d =
  Await_eff.perform (Await d) h
\end{lstlisting}

Using this we can write direct-style code using asynchronous I/O:
\begin{lstlisting}[style=ocaml]
let copy_file h src dst =
  let s = await h (File.read src) in
  await h (File.write dst s)
\end{lstlisting}

\subsection{One-shot continuations}

For various practical reasons, OCaml's algebraic effects are restricted
to only allow continuations to be resumed a single time. This rules out
interesting effects like backtracking search, but improves the ability
to reason about the interaction of effects and resources.

There is no fundamental reason that effect reflection should be
restricted to only one-shot continuations, so for the purposes of this
paper we shall ignore the restriction. However, for the same reasons as
OCaml, our actual implementation is restricted one-shot
continuations. Note that OxCaml's \lstinline[style=ocaml]{once}
mode\cite{lorenzen2024oxidizing} allows us to enforce this restriction
statically.

\subsection{Nested effect handlers}

A key feature of algebraic effects is how easy they are to compose. We
can use state whilst using asynchronous I/O by simply nesting their
handlers:
\begin{lstlisting}[style=ocaml]
handle_await (fun ch ->
  handle_state i (fun sh ->
    List.iter files (fun fl ->
      let s = await ch (File.read fl) in
      let i = int_of_string s in
      set sh (i + get sh));
    get sh))
\end{lstlisting}

This relies on the fact that the \lstinline[style=ocaml]{handle_state}
uses \lstinline[style=ocaml]{reify_local} and so is able to operate on a
local computation that closes over the local handler value
\lstinline[style=ocaml]{ch}.

If we tried to do something similar using our
\lstinline[style=ocaml]{handle_gen} handler, then we get and error:
\begin{lstlisting}[style=ocaml]
handle_await (fun ch ->
  handle_gen i (fun gh ->
    List.iter files (fun fl ->
      let s = await ch (File.read fl) in
      let i = int_of_string s in
      gen gh i)))
(*@ \textcolor{red}{Error: let s = await ch (File.read fl) in} @*)
                     (*@ \textcolor{red}{\textasciicircum\textasciicircum} @*)
(*@ \textcolor{red}{ This value is local and would escape its region.} @*)
\end{lstlisting}

We could change the definition of \lstinline[style=ocaml]{handle_gen} to
use \lstinline[style=ocaml]{reify_local} instead of
\lstinline[style=ocaml]{reify}, but then we would lose the property that
the generator we produce can be passed around freely. Instead we need to
build a handler that can forward the \lstinline[style=ocaml]{Await}
operation to different effect handlers at different points in the
program's execution.

\subsection{Effect sums}

Composing handlers, like \lstinline[style=ocaml]{handle_gen}, that use
\lstinline[style=ocaml]{reify} rather than
\lstinline[style=ocaml]{reify_local} is a fundamentally more involved
operation. Consider the case of generators that are able to use
asynchronous I/O. We are given a computation that performs both
\lstinline[style=ocaml]{Await} and \lstinline[style=ocaml]{Gen}
operations and we'd like to turn it into a stream of integers.  We would
like to be able to freely pass this stream around. In particular, we do
not wish the stream to be prevented from escaping the scope of the
current \lstinline[style=ocaml]{Await} handler.  However forcing parts
of the stream requires performing additional
\lstinline[style=ocaml]{Await} operations. Whenever we force part of the
stream we'll need to provide a handler for those
\lstinline[style=ocaml]{Await} operations.

To acheive this more general form of composition, we require an
additional capability in our effect reflection API. We need to be able
to manipulate \lstinline[style=ocaml]{Handler} types for sums of
effects.
\begin{lstlisting}[style=ocaml]
module Sum (L : Op) (R : Op) : sig
  type 'a t =
    | Left : 'a L.t -> 'a t
    | Right : 'a R.t -> 'a t
end

module Project(L : Op) (R : Op) : sig
  val outl : Handler(Sum(L)(R)).t -> Handler(L).t

  val outr : Handler(Sum(L)(R)).t -> Handler(R).t
end
\end{lstlisting}

\lstinline[style=ocaml]{Sum(L)(R)} is the coproduct of the effects
\lstinline[style=ocaml]{L} and \lstinline[style=ocaml]{R}.

Terms in the free algebra of the sum of two effects
(i.e. \lstinline[style=ocaml]{Term(Sum(L)(R)).t}) are isomorphic to a
recursive datatype that is built from the terms of one of the effects
and the operations of the other\cite{??}. We can definet this type as:
\begin{lstlisting}[style=ocaml]
module Half_term (L : Op) (R : Op) : sig
  type 'a t =
    | Return : 'a @@ global -> 'a t
    | Op : 'r R.t @@ global * ('r -> Handler(L).t @ local -> 'a t) -> 'a t
end
\end{lstlisting}
and expose the isomorphism between
\lstinline[style=ocaml]{Handler(L).t @ local -> Handler(R).t @ local -> 'a}
and \lstinline[style=ocaml]{Handler(L).t @ local -> 'a Half_term(L)(R).t}:
\begin{lstlisting}[style=ocaml]
module Half_reflection (L : Op) (R : Op) : sig
  module L_eff = Reflection(L)
  module R_eff = Reflection(R)
  module Sum_eff = Reflection(Sum(L)(R))
  module Proj = Project(L)(R)

  let reify (f : Handler(L).t @ local -> Handler(R).t @ local -> 'a) lh =
    let rec loop lh = function
      | Term(Sum(L)(R)).Return ->
          Half_term(L)(R).Return v
      | Term(Sum(L)(R)).Op(Left op, k) ->
          loop lh (k (L_eff.perform op lh))
      | Term(Sum(L)(R)).Op(Right op, k) ->
          Half_term(L)(R).Op(op, (fun res lh -> loop lh (k res)))
    in
    loop (Sum_eff.reify (fun h -> f (Proj.outl h) (Proj.outr h))) lh

  let rec reflect (f : Handler(L).t @ local -> 'a Half_term(L)(R).t) lh rh =
    match f lh with
    | Return x -> x
    | Op(op, k) -> reflect (k (R_eff.perform op rh)) lh rh
end
\end{lstlisting}

Now we can use this machinery to conveniently build a handler of
\lstinline[style=ocaml]{Gen} that also forwards operations from
\lstinline[style=ocaml]{Await}:
\begin{lstlisting}[style=ocaml]
module Await_gen_eff = Half_reflection(Await)(Gen)

type aints =
  | Finished
  | More of int * (Handler(Await).t @ local -> ints)

let handle_gen_in_await f ah =
    let rec loop = function
    | Half_term(Await)(Gen).Return () -> Finished
    | Half_term(Await)(Gen).Op(Gen i, k) -> More(i, k ())
  in
  loop (Await_gen_eff.reify f)
\end{lstlisting}
where \lstinline[style=ocaml]{handle_gen_in_await} has the type:
\begin{lstlisting}[style=ocaml]
val handle_gen_in_await :
  (Handler(Await).t @ local -> Handler(Gen).t @ local -> 'a)
  -> Handler(Await).t @ local
  -> aints
\end{lstlisting}

The literature on lexical effect handlers often ignores this issue. For
example, \cite{??} proposed a system which only supported directly
nesting effect handlers and had no mechanism for writing a composable
handler for generators that could be moved between different underlying
asynchronous I/O handlers.

This is because this composition requiring forwarding is much less
common than nesting and it fundamentally requires each use of the
\lstinline[style=ocaml]{Await} operation to perform work proportional to
the number of handlers it is forwarded through -- often referred to as a
``linear search''. There has been a tendency to treat this linear search
as some form of deficiency of algebraic effect handlers, but it is
simply a classic trade-off between the extra expressivity allowed by an
indirection and the runtime cost of traversing that indirection. With
our effect reflection API we are able to use nesting for the cases that
it supports -- avoiding the cost of the indirection -- and fall back to
forwarding for the more general cases.

\subsection{Parameterized effects and handlers}

The state handlers in this paper so far have all been restricted to
\lstinline[style=ocaml]{int} state. This is a limitation of the effect
reflection API that we have been using. We would like to support
\emph{parameterized effects} -- allowing a single handler to be used
with a whole family of effects. For example, a state handler
parameterized by the type of the state.

We can support parameterized effects by adding a module type for them,
along with parameterized versions of \lstinline[style=ocaml]{Term} and
\lstinline[style=ocaml]{Handler}:
\begin{lstlisting}[style=ocaml]
module type Op1 = sig type ('a, 'e) t end

module Term1 (O : Op1) : sig
  type ('a, 'e) t =
    | Return : 'a @@ global -> 'a t
    | Op : ('r, 'e) O.t @@ global * ('r -> 'a t) -> 'a t
end

module Handler1(O : Op1) : sig
  type 'e t
end
\end{lstlisting}
with a corresponding parameterized version of
\lstinline[style=ocaml]{Reflection}. We can then define a parameterized
state effect:
\begin{lstlisting}[style=ocaml]
module State = struct
  type ('a, 'b) t =
    | Get : 'b t
    | Set : 'b -> unit t
end
\end{lstlisting}
and write a handler that provides a parameterized interface for the
state effect:
\begin{lstlisting}[style=ocaml]
type 's handler

val handle : ('s handler @ local -> 'a) @ local -> 's -> 'a
val get : 's handler @ local -> 's
val set : 's handler @ local -> 's -> unit
\end{lstlisting}

The need for a new copy of the interface for each type constructor arity
is an artifact of OCaml's approach to higher-kinded types. Languages
with direct support for higher-kinded types should be able to support
parameterized effects of arbitrary arity with a single interface.

\section{Presheaves for algebraic effects}
\label{sec:semantics}

The design of effect reflection falls out naturally from a simple
denotational semantics built on presheaves. The key idea is to index the
meaning of types by an ambient algebraic monad of effects: rather than
have the semantics of each type be a set of elements, the semantics of
each type is a mapping from the ambient monad to a set of elements.

\subsection{Algebraic monads}

In order to index the meaning of our types by the ambient monad we need
some way of representing that ambient monad. We restrict the ambient
monad to be an \emph{algebraic monad}: one that can be expressed as the
free algebra of an algebraic theory. Such monads have nice properties,
in particular the sum of two algebraic monads always exists and is
itself an algebraic monad.

Most monads used in practice for programming are algebraic, with the
notable exception of the continuation monad. The issue with trying to
represent non-algebraic monads algebraically is essentially one of size:
if you try to write down the collection of algebirac operations for the
continuation monad you get something that is too large to be a set. This
is closely related to the strict positivity restriction on inductive
type definitions.

We recall the definitions of algebraic signature, terms of a signature,
algebraic theory and models of a theory.
\begin{definition}[Algebraic signature]
  An \emph{algebraic signature} is a set of operations with associated
  arities.
\end{definition}

\begin{example}
  The algebraic signature for boolean state is:
  \begin{equation*}
    \{ (\mathrm{get}, 2); (\mathrm{set\_true}, 1); (\mathrm{set\_false}, 1) \}
  \end{equation*}
\end{example}

\begin{definition}[Term]
  A \emph{term} $t$ of some algebraic signature $\Sigma$ is either a
  free variable $x$ or $op(t_1, \ldots, t_k)$ where $op$ is an operation
  of $\Sigma$ with arity $k$ and each $t_i$ is itself a term.
\end{definition}

\begin{definition}[Algebraic theory]
  An algebraic theory consists of an algebraic signature $\Sigma$ and a set of
  equations on terms of $\Sigma$.
\end{definition}

\begin{example}
  The algebraic theory for boolean state adds the following equations to
  the signature of boolean state:
  \begin{align*}
    &\mathrm{set\_true}(\mathrm{get}(x, y)) = \mathrm{set\_true}(x)
    &\mathrm{set\_false}(\mathrm{get}(x, y)) = \mathrm{set\_false}(y) \\
    &\mathrm{set\_true}(\mathrm{set\_true}(x)) = \mathrm{set\_true}(x)
    &\mathrm{set\_true}(\mathrm{set\_false}(x)) = \mathrm{set\_false}(x) \\
    &\mathrm{set\_false}(\mathrm{set\_true}(x)) = \mathrm{set\_true}(x)
    &\mathrm{set\_false}(\mathrm{set\_false}(x)) = \mathrm{set\_false}(x) \\
    &\mathrm{get}(\mathrm{set\_true}(x), \mathrm{set\_false}(x))) = x
  \end{align*}
\end{example}

\begin{definition}[Model]

\end{definition}

Arities in algebraic signatures are often considered as natural numbers,
but they can equivalently be just sets, with families of terms indexed
by that set used as the arguments of that operator in a term. It is also
convenient to allow families of operations parameterized by some set. As
such, we can present a signature in the form:
\begin{equation*}
\{ op_1 \types P_1 \rightarrowtail A_1; \ldots; op_n
\types P_n \rightarrowtail A_n \}
\end{equation*}
where $op \types P \rightarrowtail A$ indicates an operation $op$
parameterised by $P$ and with arity $A$.

Traditionally, algebraic signatures are restricted to the case of a
finite set of operations each with a finite arity. Most theoretical work
in the algebraic effects literature extends this to the case of
countable sets of operations with countable arities, which allows us to
construct the signature for natural number state:

\begin{example}
  The algebraic signature for natural number state is:
  \begin{equation*}
    \{ \mathrm{get} \types 1 \rightarrowtail \nat; \mathrm{set} \types \nat \rightarrowtail 1 \}
  \end{equation*}
\end{example}

However, it is not sufficient to construct the signature for state
containing a function $\nat \rightarrow \nat$. For that we must extend
algebraic signatures further to allow for arbitrary sets of operations
with arbitrary sets as arities. Note that we still require that these
collections be sets, so we continue to exclude monads like the
continuation monad.

\subsection{(Bounded) Lawvere theories}

We choose \emph{Lawvere theories} as our concrete representation of the
ambient algebraic monad. Lawvere theories are a categorical
representation of algebraic theories. However, we will mostly describe
Lawvere theories and constructions on them in terms of their
presentations as algebraic theories, so an understanding the details of
Lawvere theories is not required for understanding the rest of this
paper. We refer the interested reader to \cite{??} for a detailed
explanation of Lawvere theories.

Traditional Lawvere theories can only represent algebraic signatures
with a finite set of operations with finite arities. We instead use a
more general notion: \emph{bounded Lawvere theories}. These allow the
set of operations and their arities to be any (small) set. This allows
our effect operations to have arbitrary simple types as parameters and
results. In Appendix \ref{??}, we prove that all the properties of
Lawvere theories that we rely on can be extended from the finitary to
the bounded case.

Bounded Lawvere theories form a category $\lawc$ whose morphisms are
maps from the operations of one theory to the operations of another that
preserve the equations of the first theory according to the equations of
the second.

$\lawc$ has an initial object $\initial$, which is the algebraic theory
with no operations. It also has coproducts $L + M$ which is the theory
made by unioning the operations and equations of $L$ and $M$.

\subsection{Presheaves}

For a category $\cat$, $\psh = \setc^{\catop}$ is the category of
presheaves over $\cat$. It's objects are contravariant functors from
$\cat$ to $\setc$ and it's morphisms are natural transformations between
such functors.

A common use of presheaves is where $\cat$ has finite products and is
thought of as some form of context over which everything is
parameterized. In our case we wish to be parameterized by a context of
\emph{algebraic effects that can be mapped into the ambient monad}. Note
that the effects theories in such a context are \emph{contravariant}:
they describe effect theories that can be handled by the ambient
monad. As such, we use
\begin{equation*}
  \pshlawcop = \setc^{\lawc}
\end{equation*}
as the basis of our semantics. The $\op$ in the definition of presheaves
and the $\op$ due to our dependence being contravariant cancel out, but
it is still best to think in terms of presheaves as that is a more
precise description of what is going on here.

Our semantics will use objects from $\pshlawcop$ to represent types and
contexts, and morphisms in $\pshlawcop$ to represent terms. This allows
their meanings to be parameterized by the ambient monad. By using
presheaves rather than ordinary functions we ensure terms and types at a
particular ambient monad can always be mapped naturally to a larger
ambient monad.

Categories of presheaves have products and sums, which are defined
pointwise. We will use these as the semantics of product types and sum
types in our denotational semantics. Such categories also have
exponentials, which we will use as the semantics of functions types.

\subsection{The free algebra monad}

Given a Lawvere theory $L$, the underlying set of its free algebra forms
a monad $T_L$ on $\setc$. Given a set $A$, the values $T_L(A)$ are terms
build using the elements of $A$ and the operations of $L$ quotiented by
the equations in $L$.

\todo{Add example}

$T_{\_}$ is functorial and we can use it to construct a monad $\mon$ on
$\pshlawcop$ such that:
\begin{equation*}
\mon(P)(L) = T_L(P(L))
\end{equation*}
$\mon$ is essentially the ambient monad, and in our denotational
semantics, we will use it as our computation monad.

\subsection{The Yoneda embedding}

$\lawcop$ embeds into $\pshlawcop$ via the Yoneda embedding: for each
Lawvere theory $L$ there is a presheaf $\yoneda{L}$ in $\pshlawcop$ such
that:
\begin{align*}
&\yoneda{L}(L^{\prime}) = \homset{\lawc}{L}{L^{\prime}} \\
&\yoneda{L}(l^{\prime}) = \lambda f.\; l^{\prime} \circ f
\end{align*}
We will use $\yoneda{L}$ as the semantics for our handler types. Their
elements are just maps from $L$ into the ambient monad.

The Yoneda embedding preserves finite products, which for our
case means:
\begin{equation*}
\yoneda{L + L^{\prime}} \cong \yoneda{L} \times \yoneda{L^{\prime}}
\end{equation*}

\subsection{The global modality}

For any category $\cat$, if $\cat$ has a terminal object $\terminal$, we
can define the global modality $\glob \types [\psh, \psh]$ such that:
$\glob(P)(L) = P(\terminal)$. In the case of $\pshlawcop$ this becomes:
\begin{equation*}
\glob(P)(L) = P(\initial)
\end{equation*}

Within our semantics, this amounts to restricting a type to only the
elements which do not depend on the ambient monad. This is very similar
to the purity comonad \todo{check name} in ?? \cite{??}. In particular,
\begin{equation*}
  \glob \mon(P) \quad = \quad T_{\initial} \circ \glob
  \cong \quad \glob P
\end{equation*}
as elements of the free algebra of the initial algebraic theory are just
elements of the carrier set.

For any non-empty Lawvere theory $L$, $\glob \yoneda{L}$ is the constant
presheave that always returns the empty set because there are no
morphisms from $L$ into $\initial$. In our denotational semantics this
amounts to all non-trivial global handler types being uninhabited.

\subsection{The Yoneda lemma}

Via the Yoneda lemma, we get the following:
\begin{equation*}
(\arrow{\yoneda{L}}{\mon(\glob A)})(L') \quad \cong \quad T_{(L' + L)}(A(\terminal))
\end{equation*}
which will form the basis of our language's isomorphisms between arrows
$Y \rightarrow \glob A$ and algebraic monads $M(\glob A)$.

Exponentials in categories of presheaves are defined such that:
\begin{equation*}
\arrow{P}{Q}(L) = \homset{\pshlawcop}{\yoneda{L} \times P}{Q}
\end{equation*}

\subsection{W-types and effect sums}

\subsection{The effect reflection isomorphism}

\section{A language with effect reflection}
\label{sec:language}

We can now build a simple language with effect reflection and give it a
semantics in terms of presheaves. Our language is simply typed,
featuring all the usual type-formers. It uses a variation of
fine-grained call-by-value: it has separate syntax and typing judgements
for values and computations. The syntax is in Fig.\ref{fig:syntax} and 
the full typing rules are in Fig.\ref{fig:typing}.
\begin{figure}
\begin{align*}
  \mathrm{types}, \tau ::=\;& \tau \times \tau \mid 1 \mid \tau + \tau
                              \mid \tau \rightarrow \tau \mid
                              \ind{\Sigma}(\tau) \mid \yoneda{\Sigma}
                              \mid \glob \tau \\
  \mathrm{contexts}, \Gamma ::=\;& \cempty \mid \lbind{\Gamma}{x}{\tau} \mid \gbind{\Gamma}{x}{\tau} \\
  \mathrm{values}, v :=\;& x \mid (v, v) \mid \fst v \mid \snd v
                           \mid () \mid \inl{v} \mid \inr{v} \mid \abs{x}{\tau}{c} \\
                         &\mid \indintro{\Sigma}{op}{v}{v} \mid \ret{\Sigma}{v}
                           \mid \gbox{v} \mid \gunbox{x}{v}{v} \mid \grun{c} \\
                            &\mid \outl{v} \mid \outr{v} \\
  \mathrm{computations}, c :=\;& \return{v} \mid \letv{x}{c}{c}
                                 \mid \app{v}{v} \mid \\
                            & \case{v}{x}{c}{x}{c} \mid \\
                            & \indelim{\Sigma}{v}
                              {\indretcase{x}{c} \seq
                              \indcase{op}{x}{x}{c} \seq \ldots
                              \seq \indcase{op}{x}{x}{c}} \\
                            & \perform{v}{op}{v}{x}{c} \mid \reify{\Sigma}{x}{c}
\end{align*}
\caption{Syntax}
\label{fig:syntax}
\end{figure}

\begin{figure}
\begin{mathpar}
\inferrule{\\}
{\lbind{\Gamma}{x}{\tau} \turnv x \types \tau}
\and
\inferrule{\Gamma \turnv v_1 \types \tau_1 \\ \Gamma \turnv v_1 \types \tau_1}
{\Gamma \turnv (v_1, v_2) \types \tau_1 \times \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau_1 \times \tau_2}
{\Gamma \turnv \fst{v} \types \tau_1}
\and
\inferrule{\Gamma \turnv v \types \tau_1 \times \tau_2}
{\Gamma \turnv \snd{v} \types \tau_2}
\and
\inferrule{\\}
{\Gamma \turnv () \types 1}
\and
\inferrule{\Gamma \turnv v \types \tau_1}
{\Gamma \turnv \inl{v} \types \tau_1 + \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau_2}
{\Gamma \turnv \inr{v} \types \tau_1 + \tau_2}
\and
\inferrule{\lbind{\Gamma}{x}{\tau_1} \turnc c \types \tau_2}
{\Gamma \turnv \abs{x}{\tau_1}{c} \types \tau_1 \rightarrow \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau}
{\Gamma \turnc \return{v} \types \tau}
\and
\inferrule{\Gamma \turnc c_1 \types \tau_1 \\
           \lbind{\Gamma}{x}{\tau_1} \turnc c_2 \types \tau_2}
{\Gamma \turnc \letv{x}{c_1}{c_2} \types \tau_2}
\and
\inferrule{\Gamma \turnv v_1 \types \tau_1 \rightarrow \tau_2 \\
           \Gamma \turnv v_2 \types \tau_1}
{\Gamma \turnc v_1 v_2 \types \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau_1 + \tau_2 \\
           \lbind{\Gamma}{x_1}{\tau_1} \turnc c_1 \types \tau_3 \\
           \lbind{\Gamma}{x_2}{\tau_2} \turnc c_2 \types \tau_3}
{\Gamma \turnc \case{v}{x_1}{c_1}{x_2}{c_2} \types \tau_3}
\end{mathpar}
\caption{Standard typing rules}
\label{fig:typing-standard}
\end{figure}

\begin{figure}
\begin{mathpar}
\inferrule{\\}
{\gbind{\Gamma}{x}{\tau} \turnv x \types \tau}
\and
\inferrule{\cdiv{\Gamma} \turnv v \types \tau}
{\Gamma \turnv \gbox{v} \types \glob \tau}
\and
\inferrule{\Gamma \turnv v_1 \types \glob \tau_1 \\
           \gbind{\Gamma}{x}{\tau_1} \turnv v_2 \types \tau_2}
{\Gamma \turnv \gunbox{x}{v_1}{v_2} \types \tau_2}
\end{mathpar}
\caption{Locality typing rules}
\label{fig:typing-locality}
\end{figure}

\begin{figure}
\begin{mathpar}
\inferrule{op \types \tau_1 \rightarrowtail \tau_2 \in \Sigma \\
           \Gamma \turnv v_1 \types \glob \tau_1 \\
           \Gamma \turnv v_2 \types \glob \tau_2 \rightarrow \ind{\Sigma}(\tau)}
{\Gamma \turnv \indintro{\Sigma}{op}{v_1}{v_2} \types \ind{\Sigma}(\tau)}
\and
\inferrule{\Gamma \turnv v \types \tau}
{\Gamma \turnv \ret{\Sigma}{v} \types \ind{\Sigma}(\tau)}
\and
\inferrule{\Gamma \turnv v \types \ind{\Sigma}(\tau_1) \\
           \Sigma = \{ op_1 \types \tau_{p_1} \rightarrowtail \tau_{a_1};
             \ldots; op_n\types \tau_{p_n} \rightarrowtail \tau_{a_n} \} \\
           \lbind{\Gamma}{x_r}{\tau_1} \turnc c_r \types \tau_2 \\
           \lbind{\lbind{\Gamma}{x_{p_1}}{\glob \tau_{p_1}}}{x_{a_1}}{\glob \tau_{a_1} \rightarrow \tau_2} \turnc c_1 \types \tau_2 \\
           \ldots \\
           \lbind{\lbind{\Gamma}{x_{p_n}}{\glob \tau_{p_n}}}{x_{a_n}}{\glob \tau_{a_n} \rightarrow \tau_2} \turnc c_n \types \tau_2}
{\Gamma \turnc \indelim{\Sigma}{v}{\indretcase{x_r}{c_r} \seq
                                   \indcase{op_1}{x_{p_1}}{x_{a_1}}{c_1} \seq \ldots \seq
                                   \indcase{op_n}{x_{p_n}}{x_{a_n}}{c_n}} \types \tau_2}
\end{mathpar}
\caption{Inductive type typing rules}
\label{fig:typing-inductive}
\end{figure}

\begin{figure}
\begin{mathpar}
\inferrule{\cdiv{\Gamma} \turnc c \types \tau}
{\Gamma \turnv \grun{c} \types \tau}
\and
\inferrule{\Gamma \turnv v \types \yoneda{\Sigma_1 + \Sigma_2}}
{\Gamma \turnv \outl{v} \types \yoneda{\Sigma_1}}
\and
\inferrule{\Gamma \turnv v \types \yoneda{\Sigma_1 + \Sigma_2}}
{\Gamma \turnv \outr{v} \types \yoneda{\Sigma_2}}
\and
\inferrule{\Gamma \turnv v_1 \types \yoneda{\Sigma} \\
           op \types \tau_1 \rightarrowtail \tau_2 \in \Sigma \\
           \Gamma \turnv v_2 \types \glob \tau_1 \\
           \lbind{\Gamma}{x}{\glob \tau_2} \turnc c \types \tau_3}
{\Gamma \turnc \; \perform{v_1}{op}{v_2}{x}{c} \types \tau_3}
\and
\inferrule{\lbind{\Gamma}{x}{\yoneda{\Sigma}} \turnc c \types \glob \tau}
{\Gamma \turnc \; \reify{\Sigma}{x}{c} \types \ind{\Sigma}(\glob \tau)}
\end{mathpar}
\caption{Reflection typing rules}
\label{fig:typing-reflection}
\end{figure}

As with other algebraic effects systems in the literature, we only
support handlers for theories without any equations, even though our
semantics is defined in terms of arbitrary bounded Lawvere theories. We
define algebraic signatures $\Sigma$ as finite sets of operations
\[
\{ op_1 \types \tau_{p_1} \rightarrowtail \tau_{a_1}; \ldots; op_n
\types \tau_{p_n} \rightarrowtail \tau_{a_n} \}
\]
An operation type $\tau_p \rightarrowtail \tau_a$ indicates an operation
parameterised by $\tau_p$ and with arity $\tau_a$. Such signatures
correspond to bounded Lawvere theories without equations.

The language provides inductive types $\ind{\Sigma}$ to directly
represent terms of an algebraic signature $\Sigma$. $\ind{\Sigma}(\tau)$
has a constructor $op$ with arguments of type $\tau_p$ and
$\tau_a \rightarrow \ind{\Sigma}(\tau)$ for each operation
$op \types \tau_p \rightarrowtail \tau_a$ in $\Sigma$, as well as a
constructor $ret$ with an argument of type $\tau$. $\ind{\Sigma}$ is
sometimes called the \emph{freer monad} of $\Sigma$ \cite{??}.

The language provides types $\yoneda{\Sigma}$ to represent morphisms
from $\sem{\Sigma}$ into the ambient monad. Using these, our desired
isomorphism appears in the language as two operations, reflection
$\reflectraw{\Sigma}{op}$ and reification $\reifyraw{\Sigma}$.
\begin{mathpar}
\inferrule{\Gamma \turnv v_1 \types \yoneda{\Sigma} \\
           op \types \tau_1 \rightarrowtail \tau_2 \in \Sigma \\
           \Gamma \turnv v_2 \types \glob \tau_1 \\
           \lbind{\Gamma}{x}{\glob \tau_2} \turnc c \types \tau_3}
{\Gamma \turnc \; \perform{v_1}{op}{v_2}{x}{c} \types \tau_3}
\and
\inferrule{\lbind{\Gamma}{x}{\yoneda{\Sigma}} \turnc c \types \glob \tau}
{\Gamma \turnc \; \reify{\Sigma}{x}{c} \types \ind{\Sigma}(\glob \tau)}
\end{mathpar}

Note that we treat reflection of individual operations
$\performraw{v}{op}$ as primitive. We can define reflection of
terms $\reflectraw{\Sigma}$ in terms of $\performraw{v}{op}$:
\[
    \reflect{\Sigma}{v}{c} := \todo{write this}
\]
which has this admissible typing rule:
\begin{mathpar}
\inferrule{\Gamma \turnv v \types \yoneda{\Sigma} \\
           \Gamma \turnv c \types \ind{\Sigma}(\glob \tau)}
{\Gamma \turnc \; \reflect{\Sigma}{v}{c} \types \glob \tau}
\end{mathpar}

Our denotational semantics in terms of $\pshlawcop$ is
straightforward. We write $\sem{\Sigma}$ for the Lawvere theory
corresponding to $\Sigma$, and $F_\Sigma$ for the endofunctor on $\setc$
for which $T_{\sem{\Sigma}}$ is the free monad. For types, the
translation $\sem{\tau} \mathrel{:} \pshlawcop$ is as follows:
\begin{mathpar}
\sem{\yoneda{\Sigma}} = \yoneda{\sem{\Sigma}} \and
\sem{\glob \tau} = \glob \sem{\tau} \and
\sem{\tau_1 \rightarrow \tau_2} = \arrow{\sem{\tau_1}}{\mon(\sem{\tau_2})} \and
\sem{\ind{\Sigma}(\tau)} = \mu X.(\sem{\tau} + F_\Sigma(\mon(X)))
\end{mathpar}
where $\mu X. F(X).$ represents the initial algebra of a suitable
endofunctor. For any Lawvere theory $L$,
$\mu X.T_L(\sem{\tau} + F_\Sigma(X))$ is isomorphic to to
$T_{(L + \sem{\Sigma})}$ \cite{hyland2006combining}. That means that
$\mon(\sem{\ind{\Sigma}(\glob \tau)})$ is isomorphic to the presheaf that
maps $L$ to $T_{(L + \sem{\Sigma})}(\sem{\tau}(\terminal))$, which in
turn is isomorphic to $\sem{\yoneda{\Sigma} \rightarrow \glob \tau}$ by the
result in the previous section. This isomorphism directly provides the
semantics of reflection and reification.

\begin{align*}
&\sem{\Gamma \turnv v \types \tau} \mathrel{:} \sem{\Gamma} \rightarrow \sem{\tau} \\
&\sem{\glob \Gamma \turnv \gbox{v} \types \glob \tau} =
  \glob \sem{\cdiv{\Gamma} \turnv v \types \tau} \circ \todo{weakening and globalizing}\\
&\sem{\Gamma \turnv \gunbox{x}{v_1}{v_2} \types \tau_2} =
  \sem{\gbind{\Gamma}{x}{\tau_1} \turnv v_2 \types \tau_2}
  \circ \langle id_{\sem{\Gamma}}, \sem{\Gamma \turnv v_1 \types \glob \tau_1} \rangle
&\sem{\glob \Gamma \turnv \grun{c} \types \glob \tau} =
  \epsilon \circ \todo{removing identity monad} \circ
  \glob \sem{\cdiv{\Gamma} \turnv v \types \tau} \circ \todo{weakening and globalizing}\\
\end{align*}
\begin{align*}
&\sem{\Gamma \turnc c \types \tau} \mathrel{:} \sem{\Gamma} \rightarrow \mon(\sem{\tau}) \\
&\sem{\Gamma \turnc \return{v} \types \tau} = \eta_{\sem{\tau}} \circ \sem{\Gamma \turnv v \types \tau} \\
&\sem{\Gamma \turnc \letv{x}{c_1}{c_2} \types \tau_2} = \mu_{\sem{\tau_2}} \circ \mon(\sem{\lbind{\Gamma}{x}{\tau_1} \turnc c_2 \types \tau_2}) \circ t_{\sem{\Gamma}, \sem{\tau_1}} \circ \langle id_{\sem{\Gamma}}, \sem{\Gamma \turnc c_1 \types \tau} \rangle \\
&\sem{\Gamma \turnc \perform{v_1}{op}{v_2}{x}{c} \types \glob \tau} = \todo{fix}
  \evmap \langle \reflectc{\Sigma}{\tau} \circ \eta \circ
                   \sem{\Gamma \turnv v_1 \types \ind{\Sigma}(\glob \tau)},\,
                 \sem{\Gamma \turnv v_2 \types \yoneda{\Sigma}} \rangle \\
&\sem{\Gamma \turnc \reify{\Sigma}{x}{c} \types \ind{\Sigma}(\glob \tau)} =
  \reifyc{\Sigma}{\tau} \circ \currymap(\sem{\lbind{\Gamma}{x}{\yoneda{\Sigma}} \turnc c \types \glob \tau}) \\
\end{align*}
where $\eta$, $\mu$ and $t$ are the unit, multiplication and strength
respectively of $\mon$; $\epsilon$ is the counit of $\glob$; $\evmap$
and $\currymap$ are the evaluation map and currying map of exponentials;
$\reflectc{\Sigma}{}$ and $\reifyc{\Sigma}{}$ are the natural
transformations in the isomorphism from the previous section.

\section{Operational semantics}
\label{sec:operational}

\begin{figure}
\begin{align*}
  \mathrm{neutral\ values}, \eta :=\;& x \mid \fst \eta \mid \snd \eta
                                     \mid \gunbox{x}{\eta}{\omega} \mid \grun{\mu} \\
  \mathrm{neutral\ computations}, \mu :=\;& \letv{x}{\mu}{\gamma}
                                           \mid \app{\eta}{\omega} \mid \\
                                   & \case{\eta}{x}{\gamma}{x}{\gamma} \mid \\
                            & \indelim{\Sigma}{\eta}
                              {\indretcase{x}{\gamma} \seq
                               \indcase{op}{x}{x}{\gamma} \seq \ldots \seq
                               \indcase{op}{x}{x}{\gamma}} \\
                                   & \reify{\Sigma}{x}{\mu} \\
  \mathrm{normal\ values}, \omega :=\;& \eta \mid (\omega, \omega)
                           \mid () \mid \inl{\omega} \mid \inr{\omega} \mid \abs{x}{\tau}{\gamma} \mid \\
                           &\indintro{\Sigma}{op}{\omega}{\omega} \mid \ret{\Sigma}{\omega} \mid \\
                           &\gbox{\omega} \mid \outl{\omega} \mid \outr{\omega} \\
  \mathrm{normal\ computations}, \gamma :=\;& \mu \mid \return{\omega}
                                               \mid \perform{\eta}{op}{\omega}{x}{\gamma}
\end{align*}
\caption{Normal forms}
\label{fig:normal}
\end{figure}

\begin{figure}
\begin{align*}
\fst{(v_1, v_2)} &\reducesto v_1 \\
\snd{(v_1, v_2)} &\reducesto v_2 \\
\gunbox{x}{\gbox{v_1}}{v_2} &\reducesto \sbst{v_2}{x}{v_1} \\
\grun{\return{v}} &\reducesto v \\
\letv{x}{\return{v}}{c} &\reducesto \sbst{c}{x}{v} \\
\letv{x_1}{\perform{v_1}{op}{v_2}{x_2}{c_1}}{c_2} &\reducesto \perform{v_1}{op}{v_2}{x_2}{\letv{x_1}{c_1}{c_2}} \\
\app{(\abs{x}{\tau}{c})}{v} &\reducesto \sbst{c}{x}{v} \\
\case{(\inl{v})}{x_1}{c_1}{x_2}{c_2} &\reducesto \sbst{c_1}{x_1}{v} \\
\case{(\inr{v})}{x_1}{c_1}{x_2}{c_2} &\reducesto \sbst{c_2}{x_2}{v} \\
\indelim{\Sigma}{(\ret{\Sigma}{v})}{\indretcase{x}{c} \seq \ldots} &\reducesto \sbst{c}{x}{v} \\
\indelim{\Sigma}{\indintro{\Sigma}{op}{v_1}{v_2}}{H_1 \seq \indcase{op}{x_1}{x_2}{c} \seq H_2} \\
  \reducesto \sbst{\sbst{c}{x_1}{v_2}}{x_2}{\abs{x_a}{\tau_a}{&\letv{x_w}{\app{v_2}{x_a}}{\indelim{\Sigma}{x_w}{H_1 \seq \indcase{op}{x_1}{x_2}{c} \seq H_2}}}} \\
\reify{\Sigma}{x}{\return{v}} &\reducesto \ret{\Sigma}{v} \\
\reify{\Sigma}{x_1}{\perform{x_1}{op}{v}{x_2}{c}} &\reducesto \indintro{\Sigma}{op}{v}{\abs{x_2}{\reify{\Sigma}{x_1}{c}}} \\
\reify{\Sigma}{x_1}{\perform{x_2}{op}{v}{x_3}{c}} &\reducesto \perform{x_2}{op}{v}{x_3}{\reify{\Sigma}{x_1}{c}} \qquad (x_1 \neq x_2) \\
\perform{(\outl{v_1})}{op}{v_2}{x}{c} &\reducesto \perform{v_1}{\mathrm{left}(op)}{v_2}{x}{c} \\
\perform{(\outr{v_1})}{op}{v_2}{x}{c} &\reducesto \perform{v_1}{\mathrm{right}(op)}{v_2}{x}{c}
\end{align*}
\caption{Reduction relation}
\label{fig:reduction}
\end{figure}

\section{Implementation}
\label{sec:implementation}

\subsection{Implementation with untyped effect handlers}

We can implement the interface from Section \ref{sec:interface} directly
in terms of OCaml's untyped effect
handlers\cite{Sivaramakrishnan2021retrofitting}. The trick is to
generate a fresh effect constructor for each reification, pass that
constructor around as part of the \lstinline[style=ocaml]{handler}, and
then use it to perform operations during reflection.

\begin{lstlisting}[style=ocaml]
module Handler (O : Op) = struct
  type t = { perform : 'r. 'r O.t -> 'r }
end

module Reflection (O : Op) = struct

  let reify f =
    let module Handler =
      struct type 'a Effect.t += H : 'a O.t -> 'a Effect.t end
    in
    let handler : Handler(O).t =
      { perform = fun o -> Effect.perform (Handler.H o) }
    in
    match f handler with
    | v -> Return v
    | effect Handler.H o, k -> Op(o, fun x -> Effect.continue k x)

  let reflect t =
    fun {perform} ->
      let rec loop = function
        | Return v -> v
        | Op(o, f) -> loop (f (perform o))
      in
      loop t

  let reify_local f = exclave
    let module Handler =
      struct type 'a Effect.t += H : 'a O.t -> 'a Effect.t end
    in
    let handler = { perform = fun o -> Effect.perform (Handler.H o) } in
    match f handler with
    | v -> Return v
    | effect Handler.H o, k -> Op(o, fun x -> Effect.continue k x)

  let reflect_local t = exclave
    fun {perform} ->
      let rec loop = function
        | Return v -> v
        | Op(o, f) -> loop (f (perform o))
      in
      loop t

end
\end{lstlisting}
Sums are then implemented by just lifting the injections on operations:
\begin{lstlisting}[style=ocaml]
module Sum (L : Op) (R : Op) = struct

  module Op = struct
    type 'a t =
      | Left : 'a L.t -> 'a t
      | Right : 'a R.t -> 'a t
  end

  let outl {perform} =
    {perform = fun op -> perform (Left op)}

  let outr {perform} =
    {perform = fun op -> perform (Right op)}

end
\end{lstlisting}

\subsection{More efficient implementation in terms of fibers}

In the above implementation, whenever an operation is reflected we
perform a linear search up the stack of handlers until we find the
corresponding one. This search is not actually necessary...

\section{Related work}
\label{sec:related-work}

\section{Future work}
\label{sec:future-work}

\bibliography{references}
\bibliographystyle{plain}

\section{Appendix}

Previous theoretical work on algebraic effects, such as \cite{??},
has tended to use \emph{countable Lawvere theories}. These correspond to
monads who rank is \todo{how should I write the countable
  cardinality}. This still restricts the parameters and arity operations
to be countable. We instead use \emph{bounded Lawvere
  theories}, corresponding to monads that have some cardinal as their
rank. This allows parameters and arities of operations to be arbitrary
simple types.

Bounded Lawvere theories are not given a consistent name in the
literature. A number of results that are proved for countable Lawvere
theories are asserted to be true for the bounded case. As we could not
find actual proofs for these claims we have included such proofs in
Appendix \ref{??}. We also include a proof of an interesting property
of bounded Lawvere theories:

\todo{How to format and label this?}  {\b Bounded Lawvere theories
  correspond precisely to those (infinitary) Lawvere theories whose
  term monad can be expressed as a W-type}

\end{document}